Consistency and variability in the development of children's drawings of object categories

### 2021-07-07 writing/Manuscript_June2021.Rmd
Latest edits to intro following MCF's comments

---
title             : "Consistency and variability in the development of children's drawings of object categories"
shorttitle        : "Development of drawing"

@@

In modern times, drawings are produced prolifically by children from an early age. 
Figurative drawings have long provided inspiration for scientists investigating children's emerging cognitive abilities [@minsky1972artificial], and accordingly a long history of work has examined changes in children's drawings across development [@piaget1929child; @kellogg1969analyzing; @karmiloff1990constraints; @fury1997children].
Indeed, there appear to be dramatic changes in how children encode diagnostic visual information in their drawings across age; younger children (4-5 years) tend to include fewer cues in their drawings to differentiate between target concepts (e.g., \textit{adult} vs. \textit{child}) than older children, who enrich their drawings with more diagnostic part [@sitton1992drawing] and relational [@light1983effects] information. 
Overall, younger children produce drawings of object categories that are less diagnostic of the categories they are trying to depict [@long2018drawings; @long2021parallel].

What drives these dramatic changes in children's drawings across development? 
A common view is that these changes are driven primarily by children's increasing ability to plan and control their motor movements [@freeman1987current; @rehrig2018does]. 
While such changes in visuomotor control are clearly important, this view fails to account for other important constraints, such as how well children are able to access previously acquired semantic knowledge about each target concept and maintain this information in working memory during drawing production.
Thus, drawing is a skill that likely also relies on children's evolving perceptual category representations [@long2018drawings; @natu2016development; @dekker2011dorsal] as well as their increasing working memory capacity [@pailian2016visual].

One idea is that a principal reason younger children produce less recognizable drawings is because they simply have more difficulty recalling the relevant perceptual features of different categories: that is, when asked to "draw a [rabbit]", they may struggle to conjure up the relevant perceptual details and then hold in mind what rabbits tend to look like. 
On this account, providing children with additional perceptual information about different categories -- for example, canonical photographs of typical exemplars -- could help them improve their drawings of these categories, as it does with adults [@fan2020relating]. 
However, prior work also suggests that younger children tend to draw what they know about objects rather than integrate the information in their immediate perceptual experience.  
For example, when asked to draw from observation, younger children tend to include features that are not visible from their vantage point, yet are diagnostic of category membership (e.g., a handle on a \textit{mug}) [@bremmer1984prior; @barrett1976symbolism], and only omit these features later in development. 
Similarly, young children will insist that their nearly identical drawings of different concepts (e.g., balloon and person) unambiguously refer to different things [@bloom1998intention]. 
Thus, an alternative possibility is that only older children may be able to produce more recognizable drawings when provided with canonical exemplars of different categories. 
On this account, changes in children's drawings from memory across age may be largely due to other factors beyond changes in memory -- for example, changes in how children represent the diagnostic visual features of each category [@long2018drawings].

To tease apart these alternatives, we investigated the development of children's ability to produce recognizable drawings of visual concepts when asked to draw from memory ("Can you draw a [rabbit]?) and when asked to draw a canonical exemplar from observation ("Can you draw this [rabbit] as it looks in the picture?") across two different geographical sites (San Jose, USA and Beijing, China).
To do so, we gathered digital drawings from a large sample of children 4-9 years of age in a controlled experimental setting.
While children in different communities may spend more or less time practicing drawing or use different visual conventions to draw [@cohn2012explaining; @willats2006making], most empirical studies on children's drawings have been conducted exclusively on children from the United States -- or focused on differences in educational practices between communities [@la2001children; @winner1989can; @huntsinger2011cultural].
In addition, we assessed the degree to which visuomotor development accounts for developmental changes in drawing recognizability by measuring each child's shape tracing abilities.

We had several predictions about the nature of the effects that we would observe. First, we predicted that only older children would be able to use the visual information present in the canonical photographs to improve their drawings, in keeping with accounts of naïve realism and contra a strong account of working memory limitations.
Second, we predicted that we would see convergence in the development of drawing abilities across both geographical sites, with older children becoming progressively better at producing recognizable drawings.
Finally, we predicted that most of the variance across geographical sites in drawing ability would be explainable by differences in visuomotor control, operationalized as performance on a shape tracing task; these primary analyses were pre-registered at https://osf.io/qymjr/.

@@

## Participants
265 participants were recruited from the San Jose Children’s Discovery museum, the Palo Alto Junior Museum and Zoo, and preschool and elementary schools outside of Beijing; approximately equal numbers of participants were recruited in Northern California and the Beijing area. Our goal was to recruit 120 children between 4-9 years of age after exclusions (i.e. 20 4-year-olds, 20 5-year-olds, etc.) at each geographical site. In the San Jose based sample, 135 children participated in the experiment; 6 participants were excluded, (3) for skipping more than 6 drawing trials and (3) for scribbling three or more times in a row. Six additional participants were tested but their data was not recorded due to a technical error, and two participants never advanced past the practice trials, leading to a final sample of `r count_by_site$num_participants[1]` children. In the Beijing based sample, `r count_by_site$num_participants[2]` children participated; an additional 8 participants were tested but their data was not recorded due to a technical error with the remote database. Two 10-year-olds (aged 10 years, 0 months and 10 years, 1 month) were accidentally tested and included in the 9-year-old age group. On average, each child contributed `r mean(count_by_id$num_drawings)` to analysis (min `r min(count_by_id$num_drawings)`, max = `r max(count_by_id$num_drawings)`). No additional demographic data was recorded about the participants. This protocol was approved by both the Institutional Review Board at [blinded] (43992, Development of Children's Drawing Abilities) and the Department of Psychology Ethics Committee at [blinded] in Beijing, China.

@@

The primary experimental manipulation was the drawing task that children participated in; each child completed both tasks and the order of tasks was counterbalanced across participants. To elicit drawings from memory, children saw a short video clip where an experimenter said,  “What about a [cat]? Can you draw a [cat]?” (exact Mandarin translation has trouble in latex). To elicit drawings from observation, children heard audio of the same experimenter who said,  “What about this [cat]? Can you draw the [cat] as it looks in the picture?”  (Mandarin translation) while seeing a canonical photograph of each category; this photograph then remained on the screen for the duration of the drawing trial.  This photograph was randomly sampled from one of three possible exemplars. Overall, each child completed 2 tracing trials and 12 drawing trials, with six categories randomly assigned to each condition and displayed in a random order. Task examples are shown in Figure \ref{fig:example-tasks}.

@@

# Results
Overall, we found relative consistency in the developmental trajectory of children's drawings. Figure \ref{fig:main-results} shows the recognizability of children's drawings at each age as a function of the drawing task and the geographical site that they were located in; All model coefficients can be found in Table 1.  First, we found steady changes in the recognizability of children's drawings across this age range, replicating prior work with an observational dataset collected at a free-standing kiosk [@long2021parallel]. Critically, these changes in recognizability did not depend on the drawing task children completed: children's drawings were equally recognizable when they were asked to "Draw this rabbit as it looks in the picture" and they were simply asked "What about a rabbit? Can you draw a rabbit." While we anticipated that older children might produce more recognizable drawings when drawing from observation -- as do adults when presented with canonical photographs of object categories [@yang2021visual] -- this was not the case.

However, we did observe a main effect of geographical site: Children who were recruited near Beijing, China tended to produce more recognizable drawings than children who were recruited near San Jose, USA. In a second set of analyses, we aimed to understand the nature of this difference in recognizability. We hypothesized that any differences in the recognizability of children's drawings across sites would be mediated by the amount of effort children expended while drawing or their tracing abilities.  We thus first examined how effort and tracing abilities varied across task, age, and geographical location. Figure \ref{fig:effort-covariates} shows three effort covariates - average intensity, number of strokes used, and time spent drawing -- measured for each drawing as a function of children's age, drawing task, and geographical site.  Cchildren tended to spend more time drawing when asked to draw "the [rabbit] as it looks in the picture", and children recruited outside Beijing tended to spend more time drawing overall. In contrast, children's tracing abilities did not vary across geographical site: Figure \ref{fig:tracing} shows the average tracing score at each age at each site. While children's tracing abilities increased steadily with age, as in prior work [@long2021parallel], children recruited at both sites produced equally good tracings.

@@

Figure \ref{fig:item-effects} shows the recognizability of children's drawings for each of the 12 categories a function of geographical location and children's age and reveal considerable variability.  For example, certain categories (e.g., cats and rabbits) were overall more difficult for children to depict such that they were distinct from one another, particularly for younger children. In addition, children recruited near Beijing produced more recognizable drawings of certain categories at all ages -- including airplanes, birds, and rabbits -- while older children recruited near San Jose produced more recognizable drawings of bikes.  Figure \ref{fig:example-drawings} shows recognizable, randomly sampled example drawings from 6-year-olds at each category, site, and condition. These exploratory analyses thus reveal systematic differences between object categories and geographical site that do not seem easily attributable to differences in effort.

>>>>

---
title             : "Developmental consistency in children’s drawings of object categories across two countries"
shorttitle        : "Development of drawing"

@@

In modern times, drawings are produced prolifically by children from an early age. 
Figurative drawings have long provided inspiration for scientists investigating children's emerging cognitive abilities [@minsky1972artificial], and accordingly a long history of work has examined changes in children's drawings across development [@piaget1929child; @kellogg1969analyzing; @karmiloff1990constraints; @fury1997children].
Indeed, there appear to be dramatic changes in how children encode diagnostic visual information in their drawings across age; younger children (4-5 years) tend to include fewer cues in their drawings to differentiate between target concepts (e.g., \textit{adult} vs. \textit{child}) than older children, who enrich their drawings with more diagnostic part [@sitton1992drawing] and relational [@light1983effects] information.

What drives these dramatic changes in children's drawings across development? 
A common view is that these changes are driven primarily by children's increasing ability to plan and control their motor movements [@freeman1987current; @rehrig2018does]. 
While such changes in visuomotor control are clearly important, this view fails to account for other important constraints, such as how well children are able to access previously acquired semantic knowledge about each target concept and maintain this information in working memory during drawing production.
In our prior work, we found some evidence that these changes in children's drawings are accompanied by changes in children's ability to recognize drawings of object categories: in a correlational study, older children produced drawings of object categories that were more diagnostic of the categories they are trying to depict [@long2018drawings; @long2021parallel] and tended to rely more on these diagnostic visual features during recognition. 
Thus, drawing is a skill that likely also relies on children's evolving perceptual category representations [@long2018drawings; @natu2016development; @dekker2011dorsal] as well as their increasing working memory capacity [@pailian2016visual].

Here, we directly test the idea that a principal reason younger children produce less recognizable drawings is because they simply have more difficulty recalling the relevant perceptual features of different categories: that is, when asked to "draw a [rabbit]", they may struggle to conjure up the relevant perceptual details and then hold in mind what rabbits tend to look like. 
On this account, providing children with additional perceptual information about different categories -- for example, canonical photographs of typical exemplars -- could help them improve their drawings of these categories, as it does with adults [@fan2020relating]. 
However, prior work also suggests that younger children tend to draw what they know about objects rather than integrate the information in their immediate perceptual experience.  
For example, when asked to draw from observation, younger children tend to include features that are not visible from their vantage point, yet are diagnostic of category membership (e.g., a handle on a \textit{mug}) [@bremmer1984prior; @barrett1976symbolism], and only omit these features later in development. 
Similarly, young children will insist that their nearly identical drawings of different concepts (e.g., balloon and person) unambiguously refer to different things [@bloom1998intention]. 
Thus, an alternative possibility is that only older children may be able to produce more recognizable drawings when provided with canonical exemplars of different categories. 
On this account, changes in children's drawings from memory across age may be largely due to other factors beyond changes in memory -- for example, changes in how children represent the diagnostic visual features of each category [@long2018drawings].

To tease apart these alternatives, we investigated the development of children's ability to produce recognizable drawings of visual concepts when children were provided with a verbal cue ("Can you draw a [rabbit]?) versus when provided with a picture cue ("Can you draw this [rabbit] as it looks in the picture?"). On verbal-cue trials, children thus must access their mental representation of a "rabbit" and choose the features necessary to convey that object's identity. Conversely, on picture-cue trials, children are explicitly asked to rely on the visual features provided in a canonical photograph of each object category.

Second, we tested the generalizability of our findings across different populations by recruiting children from two different sites in different countries — San Jose, USA and Beijing, China.
Most empirical studies on children's drawings have been conducted exclusively on small samples of children from the United States or Western Europe, limiting their generalizability. Further, children in different communities may spend more or less time practicing drawing or use different visual conventions to draw [@cohn2012explaining; @willats2006making]. While some prior work has focused on differences in children's drawings across these two countries, most of this descriptive work has focused on differences in educational practices that may account for any observed differences [@la2001children; @winner1989can; @huntsinger2011cultural]. Instead, here we focus on the degree to which the overall developmental trajectory of drawing abilities are similar between children at these two sites.

Finally, we assessed the degree to which visuomotor development accounts for observed developmental changes in drawing recognizability. While it is clear that visuomotor control influences how and what we can draw, little work has directly related meausres of visuomotor control to drawing outcomes. To do so, we measured each child's visuomotor control via a shape tracing task on a table and related these measurements of tracing accuracy to the recognizability of the drawings that each child produced.

In the present study, we thus collect shape tracings and digital drawings of visual concepts from 4-9 year-old children in two different countries using both picture cues and verbal cues.  In doing so, we make three major contributions to our understanding of the development of children's drawings. First, we replicate our prior findings  [@long2021parallel] that the recognizability of children's drawings increases steadily throughout this age range (4-9 years). Second, we test the degree to which working memory constraints might account for these developmental changes. In keeping with accounts of naïve realism and contra a strong account of working memory limitations, we predicted that only older children would be able to use the visual information present in the canonical photographs to improve their drawings, Third, we test the generalizability of our findings across two different countries. We predicted that we would see convergence in the development of drawing abilities across both geographical sites, with older children becoming progressively better at producing recognizable drawings. In addition, we predicted that most of the variance across geographical sites in drawing ability would be explainable by differences in visuomotor control, operationalized as performance on a shape tracing task; these primary analyses were pre-registered at https://osf.io/qymjr/.

@@

## Participants
265 participants were recruited from two local children's museum in [BLINDED] and preschool and elementary schools outside of Beijing; approximately equal numbers of participants were recruited in Northern California and the Beijing area. Our goal was to recruit 120 children between 4-9 years of age after exclusions (i.e. 20 4-year-olds, 20 5-year-olds, etc.) at each geographical site. In the San Jose based sample, 135 children participated in the experiment; 6 participants were excluded, (3) for skipping more than 6 drawing trials and (3) for scribbling three or more times in a row. Six additional participants were tested but their data was not recorded due to a technical error, and two participants never advanced past the practice trials, leading to a final sample of `r count_by_site$num_participants[1]` children. In the Beijing based sample, `r count_by_site$num_participants[2]` children participated; an additional 8 participants were tested but their data was not recorded due to a technical error with the remote database. Two 10-year-olds (aged 10 years, 0 months and 10 years, 1 month) were accidentally tested and included in the 9-year-old age group. On average, each child contributed `r mean(count_by_id$num_drawings)` drawings to analysis (min `r min(count_by_id$num_drawings)`, max = `r max(count_by_id$num_drawings)`). No additional demographic data was recorded about the participants. This protocol was approved by both the Institutional Review Board at [blinded] (43992, Development of Children's Drawing Abilities) and the Department of Psychology Ethics Committee at [blinded] in Beijing, China.

@@

The primary experimental manipulation was the drawing task that children participated in; each child completed both tasks and the order of tasks was counterbalanced across participants. Overall, each child completed 2 tracing trials and 12 drawing trials, with six categories randomly assigned to each condition and displayed in a random order. on verbal-cue trials, children saw a short video clip where an experimenter said,  “What about a [cat]? Can you draw a [cat]?” (exact Mandarin translation has trouble in latex). On picture-cue trials, children heard audio of the same experimenter who said,  “What about this [cat]? Can you draw the [cat] as it looks in the picture?”  (Mandarin translation) while seeing a canonical photograph of each category; this photograph then remained on the screen for the duration of the drawing trial.  This photograph was randomly sampled from one of three possible exemplars. Task examples are shown in Figure \ref{fig:example-tasks}.

@@

# Results
### Confirmatory Analyses
Overall, we found relative consistency in the developmental trajectory of children's drawings. Figure \ref{fig:main-results} shows the recognizability of children's drawings at each age as a function of the drawing task and the geographical site that they were located in; All model coefficients can be found in Table 1.  First, we found steady changes in the recognizability of children's drawings across this age range, replicating prior work with an observational dataset collected at a free-standing kiosk [@long2021parallel]. Critically, these changes in recognizability did not depend on the drawing task children completed: children's drawings were equally recognizable when they were asked to "Draw this rabbit as it looks in the picture" and they were simply asked "What about a rabbit? Can you draw a rabbit." While we anticipated that older children might produce more recognizable drawings when drawing from observation -- as do adults when presented with canonical photographs of object categories [@yang2021visual] -- this was not the case.

However, we did observe a main effect of geographical site: Children from Beijing, China tended to produce more recognizable drawings than children from San Jose, USA. Figure \ref{fig:main-results} suggests that this may be because children from Beijing, China tended to reach
In a second set of analyses, we aimed to understand the nature of this difference in recognizability. We hypothesized that any differences in the recognizability of children's drawings across sites would be mediated by the amount of effort children expended while drawing or their tracing abilities.  We thus first examined how effort and tracing abilities varied across task, age, and geographical location. Figure \ref{fig:effort-covariates} shows three effort covariates - average intensity, number of strokes used, and time spent drawing -- measured for each drawing as a function of children's age, drawing task, and geographical site.  Cchildren tended to spend more time drawing when asked to draw "the [rabbit] as it looks in the picture", and children recruited outside Beijing tended to spend more time drawing overall. In contrast, children's tracing abilities did not vary across geographical site: Figure \ref{fig:tracing} shows the average tracing score at each age at each site. While children's tracing abilities increased steadily with age, as in prior work [@long2021parallel], children recruited at both sites produced equally good tracings.

@@

### Exploratory Analyses
Figure \ref{fig:item-effects} shows the recognizability of children's drawings for each of the 12 categories a function of geographical location and children's age and reveal considerable variability.  For example, certain categories (e.g., cats and rabbits) were overall more difficult for children to depict such that they were distinct from one another, particularly for younger children. In addition, children recruited near Beijing produced more recognizable drawings of certain categories at all ages -- including airplanes, birds, and rabbits -- while older children recruited near San Jose produced more recognizable drawings of bikes.  Figure \ref{fig:example-drawings} shows recognizable, randomly sampled example drawings from 6-year-olds at each category, site, and condition. These exploratory analyses thus reveal systematic differences between object categories and geographical site that do not seem easily attributable to differences in effort.
